---
title: "Abundant species dynamics from benthic surveys"
author: "An Bui"
date: "1/31/2022"
output: 
  html_document:
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: no
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0. set up
```{r warning = FALSE, message = FALSE}
# source the cleaning/summarizing script, which sources the set up
source(here::here("code", "01_cleaning-and-summarizing.R"))
```

# 1. Biomass

## a. Guiding questions

What's the percent community biomass that can be attributed to:  
1. _Laminariales_, and  
2. the 11 abundant species from Miller et al.?  

Data question: what is the difference between the percent cover data and the percent cover that's in the column in the biomass data?  

Stray thoughts: introduced/native species trait column?

## b. Test case: SCTW from 3 different sampling points

To make sure the code was working, I tested this out with "SCTW_2010-07-29", "SCTW_2016-07-20", and "SCTW_2020-07-29". The issue comes from scenarios where species were counted at more than one transect - for example, at SCTW_2020-07-29, EC was at both transect 2 and 3, but I only want the dry/wet biomass for EC for the whole survey.  

When I create the data frame, there are then "duplicate" observations for total species biomass for EC because it was present at both transects. To fix this, I filtered the data frame for unique observations of the total and percent biomass calculations and double checked the total dry percent to make sure everything added up to 1.  

### i. Calculations

This is just a test to make sure the calculations worked.

```{r}
test <- biomass %>% 
  filter(sample_ID %in% c("SCTW_2010-07-29", "SCTW_2016-07-20", "SCTW_2020-07-29")) %>% 
  select(sample_ID, site, year, date, sp_code, dry_gm2, wm_gm2) %>% 
  # create a new column where total biomass for both surveys is calculated
  group_by(site, year) %>% 
  mutate(total_dry = sum(dry_gm2),
         total_wet = sum(wm_gm2)) %>% 
  ungroup() %>% 
  # create a new column where total biomass for the species for the whole survey is calculated
  group_by(site, year, sp_code) %>% 
  mutate(total_sp_dry = sum(dry_gm2),
         total_sp_wet = sum(wm_gm2)) %>% 
  ungroup() %>% 
  # create a new column where percent of total biomass per survey for each species is calculated
  mutate(percent_sp_dry = total_sp_dry/total_dry,
         percent_sp_wet = total_sp_wet/total_wet) %>% 
  # double check to make sure the percents worked out...
  select(sample_ID, site, year, date, sp_code, 
         # values that need to be unique: 
         total_sp_dry, total_sp_wet, percent_sp_dry, percent_sp_wet) %>% 
  unique() %>% 
  group_by(sample_ID) %>% 
  mutate(total_percent = sum(percent_sp_dry),
         total_percent_wet = sum(percent_sp_wet)) %>% 
  # join with coarse traits data frame, which includes taxonomy
  left_join(., coarse_traits, by = "sp_code")
```

### ii. Plots

These are tests of the plots.

```{r}
# plot of proportion total biomass by phylum faceted by sampling date
ggplot(test) +
  aes(x = taxon_phylum, y = percent_sp_dry) +
  # calculating sum of proportion within stat_summary
  stat_summary(aes(fill = taxon_phylum),
               fun = "sum", geom = "bar") +
  # adding labels for sum values
  stat_summary(aes(label = round(..y.., 2)), 
               fun = "sum", geom = "text", 
               size = 4, vjust = -0.5) +
  scale_fill_manual(values = c("Chlorophyta" = chloro_col, 
                               "Ochrophyta" = ochro_col, 
                               "Rhodophyta" = rhodo_col)) +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  facet_wrap(~sample_ID) +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "Phylum",
       y = "Proportion total biomass",
       title = "Biomass by phylum for SCTW")

# pretend timeseries
ggplot(test) +
  aes(x = date, y = percent_sp_dry) +
  # calculating sum of proportion within stat_summary
  stat_summary(aes(col = taxon_phylum),
               fun = "sum", geom = "line", size = 1) +
  stat_summary(aes(col = taxon_phylum),
               fun = "sum", geom = "point", size = 2) +
  scale_color_manual(values = c("Chlorophyta" = chloro_col, 
                                "Ochrophyta" = ochro_col, 
                                "Rhodophyta" = rhodo_col)) +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  theme_bw() +
  # theme(legend.position = "none") +
  labs(x = "Sampling date",
       y = "Proportion total biomass",
       title = "Biomass by phylum through time for SCTW",
       col = "Phylum")
```

## c. Biomass data set

Since this worked, I felt comfortable doing this for the whole `biomass` data frame.  

### i. Calculations

**Note:** as of 2022-03-04, the code to create the `prop_biomass` data frame is in the `00_set-up.R` script.  

### ii. Plots

Made some plot functions to keep things tidy.  

**Note:** FIX THIS. YOU NEED TO REDO THESE PLOTS BECAUSE THE DATA FRAME THEY'RE BUILT ON DOESN'T TAKE THE AVERAGE AT THE SITE ANYMORE.  
```{r}
phylum.plot <- function(site_choice) {
  # choose a subtitle from sites_full
  subtitle <- pluck(sites_full, site_choice)
  
  ggplot(prop_biomass %>% filter(site == {{ site_choice }})) +
  aes(x = date, y = percent_sp_dry) +
  # calculating sum of proportion within stat_summary
  stat_summary(aes(col = taxon_phylum),
               fun = "sum", geom = "line", size = 1) +
  stat_summary(aes(col = taxon_phylum),
               fun = "sum", geom = "point", size = 2) +
  scale_color_manual(values = c("Chlorophyta" = chloro_col, 
                                "Ochrophyta" = ochro_col, 
                                "Rhodophyta" = rhodo_col)) +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  theme_bw() +
  # theme(legend.position = "none") +
  labs(x = "Sampling date",
       y = "Proportion total biomass",
       title = "Biomass by phylum through time",
       subtitle = subtitle,
       col = "Phylum")
}

growth.form.plot <- function(site_choice) {
  # choose a subtitle from sites_full
  subtitle <- pluck(sites_full, site_choice)
  
  ggplot(prop_biomass %>% filter(site == {{ site_choice }})) +
    aes(x = date, y = percent_sp_dry) +
    # calculating sum of proportion within stat_summary
    stat_summary(aes(col = growth_form),
                 fun = "sum", geom = "line", size = 1) +
    stat_summary(aes(col = growth_form),
                 fun = "sum", geom = "point", size = 2) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
    theme_bw() +
    labs(x = "Sampling date",
         y = "Proportion total biomass",
         title = "Biomass by growth form through time",
         subtitle = subtitle,
         col = "Growth form")
}

spp.interest.plot <- function(site_choice) {
  # choose a subtitle from sites_full
  subtitle <- pluck(sites_full, site_choice)
  
  ggplot(prop_biomass %>% filter(site == {{ site_choice }})) +
    aes(x = date, y = percent_sp_dry) +
    # calculating sum of proportion within stat_summary
    stat_summary(aes(col = algae_interest),
                 fun = "sum", geom = "line", size = 1) +
    stat_summary(aes(col = algae_interest),
                 fun = "sum", geom = "point", size = 2) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
    theme_bw() +
    labs(x = "Sampling date",
         y = "Proportion total biomass",
         title = "Species of interest",
         subtitle = subtitle, 
         col = " ")
}

species.plot <- function(site_choice) {
  # choose a title from sites_full
  title <- pluck(sites_full, site_choice)
  
  df <- prop_biomass %>% ungroup() %>% filter(site == {{ site_choice }})
  
  plot_ly(data = df, 
          x = ~date, y = ~percent_sp_dry, 
          color = ~scientific_name, colors = c(cal_palette(name = "tidepool", type = "continuous")), 
          mode = "markers+lines") %>% 
    add_trace(mode = "markers+lines", type = "scatter") %>% 
    layout(title = title, 
           xaxis = list(title = "Date", nticks = 8), 
           yaxis = list(title = "Proportion total biomass", range = c(-0.01, 1)),
           legend = list(title = list(text = '<b> Scientific name </b>')))
}

```

```{r}
phylum.plot("bull")
phylum.plot("aque")
phylum.plot("ahnd")
phylum.plot("napl")
phylum.plot("ivee")
phylum.plot("golb")
phylum.plot("abur")
phylum.plot("mohk")
phylum.plot("carp")
phylum.plot("scdi")
phylum.plot("sctw")

growth.form.plot("bull")
growth.form.plot("aque")
growth.form.plot("ahnd")
growth.form.plot("napl")
growth.form.plot("ivee")
growth.form.plot("golb")
growth.form.plot("abur")
growth.form.plot("mohk")
growth.form.plot("carp")
growth.form.plot("scdi")
growth.form.plot("sctw")

spp.interest.plot("bull")
spp.interest.plot("aque")
spp.interest.plot("ahnd")
spp.interest.plot("napl")
spp.interest.plot("ivee")
spp.interest.plot("golb")
spp.interest.plot("abur")
spp.interest.plot("mohk")
spp.interest.plot("carp")
spp.interest.plot("scdi")
spp.interest.plot("sctw")

# these use plotly so that you can look at species on their own
species.plot("bull")
species.plot("aque")
species.plot("ahnd")
species.plot("napl")
species.plot("ivee")
species.plot("golb")
species.plot("abur")
species.plot("mohk")
species.plot("carp")
species.plot("scdi")
species.plot("sctw")
```

# 2. Tables of species contribution to biomass

This is a big table of percent species contribution to community biomass for each site across all sampling points. It's useful for figuring out where to sample what. `05_species-selection-tables.R` is a separate script with this code, essentially to play around with sites.

```{r}
table_df <- biomass %>% 
  select(site, scientific_name, dry_gm2, wm_gm2) %>% 
  mutate(site = fct_relevel(site, "bull", "aque", "ahnd", "napl", "ivee", "golb", "abur",
                            "mohk", "carp", "scdi", "sctw")) %>% 
  # filter out MAPY
  filter(scientific_name != "Macrocystis pyrifera") %>% 
  # create a new column where total biomass for the whole survey is calculated
  group_by(site) %>% 
  mutate(total_dry = sum(dry_gm2),
         total_wet = sum(wm_gm2)) %>% 
  ungroup() %>% 
  # create a new column where total biomass for the species for the whole survey is calculated
  group_by(site, scientific_name) %>% 
  mutate(total_sp_dry = sum(dry_gm2),
         total_sp_wet = sum(wm_gm2)) %>% 
  ungroup() %>% 
  # create a new column where percent of total biomass per survey for each species is calculated
  mutate(percent_sp_dry = total_sp_dry/total_dry,
         percent_sp_wet = total_sp_wet/total_wet) %>% 
  # double check to make sure the percents worked out...
  select(site, scientific_name, 
         # values that need to be unique: 
         total_sp_dry, total_sp_wet, percent_sp_dry, percent_sp_wet) %>% 
  unique() %>% 
  group_by(site) %>% 
  mutate(total_percent = sum(percent_sp_dry),
         total_percent_wet = sum(percent_sp_wet)) %>% 
  select(scientific_name, site, percent_sp_dry) %>% 
  mutate(percent_sp_dry = round(percent_sp_dry, 4)) %>% 
  mutate(percent_whole_sp_dry = percent_sp_dry*100) %>% 
  select(-percent_sp_dry) %>% 
  mutate(site = str_to_upper(site)) %>% 
  pivot_wider(names_from = "site", values_from = "percent_whole_sp_dry") %>% 
  left_join(., coarse_traits, by = "scientific_name") %>% 
  slice(-2) %>% 
  select(scientific_name, sp_code, BULL, AQUE, AHND, NAPL, IVEE, GOLB, ABUR,
                            MOHK, CARP, SCDI, SCTW)

table <- table_df %>% 
  gt() %>% 
  data_color(columns = 3,
             colors = scales::col_numeric(palette = gradient_palette, 
                                          domain = c(max(table_df[, 3]), 0))) %>% 
  data_color(columns = 4,
             colors = scales::col_numeric(palette = gradient_palette, 
                                          domain = c(max(table_df[, 4]), 0))) %>% 
  data_color(columns = 5,
             colors = scales::col_numeric(palette = gradient_palette, 
                                          domain = c(max(table_df[, 5]), 0))) %>% 
  data_color(columns = 6,
             colors = scales::col_numeric(palette = gradient_palette, 
                                          domain = c(max(table_df[, 6]), 0))) %>% 
  data_color(columns = 7,
             colors = scales::col_numeric(palette = gradient_palette, 
                                          domain = c(max(table_df[, 7]), 0))) %>% 
  data_color(columns = 8,
             colors = scales::col_numeric(palette = gradient_palette, 
                                          domain = c(max(table_df[, 8]), 0))) %>% 
  data_color(columns = 9,
             colors = scales::col_numeric(palette = gradient_palette, 
                                          domain = c(max(table_df[, 9]), 0))) %>% 
  data_color(columns = 10,
             colors = scales::col_numeric(palette = gradient_palette, 
                                          domain = c(max(table_df[, 10]), 0))) %>% 
  data_color(columns = 11,
             colors = scales::col_numeric(palette = gradient_palette, 
                                          domain = c(max(table_df[, 11]), 0))) %>% 
  data_color(columns = 12,
             colors = scales::col_numeric(palette = gradient_palette, 
                                          domain = c(max(table_df[, 12]), 0))) %>% 
  data_color(columns = 13,
             colors = scales::col_numeric(palette = gradient_palette, 
                                          domain = c(max(table_df[, 13]), 0))) %>% 
  fmt_missing(columns = 3:13,
              missing_text = "--") %>% 
  tab_spanner(
    label = "Sites",
    columns = c(BULL, AQUE, AHND, NAPL, IVEE, GOLB, ABUR,
                MOHK, CARP, SCDI, SCTW)
  ) %>% 
  tab_spanner(
    label = "Species",
    columns = c("scientific_name", "sp_code")
  ) %>% 
  cols_label(
    scientific_name = "Scientific name",
    sp_code = "Species code",
  )
#   %>% gtsave("sp_site_table-selection.png", path = here::here("tables"))
```

```{r}
filtered_table_df <- biomass %>% 
  select(site, scientific_name, dry_gm2, wm_gm2) %>% 
  mutate(site = fct_relevel(site, "bull", "aque", "ahnd", "napl", "ivee", "golb", "abur",
                            "mohk", "carp", "scdi", "sctw")) %>% 
  mutate(spp_interest = case_when(
    scientific_name %in% c("Desmarestia ligulata", 
                   "Stephanocystis osmundacea", 
                   "Sargassum horneri", 
                   "Sargassum muticum", 
                   "Eisenia arborea", 
                   "Egregia menziesii", 
                   "Laminaria farlowii", 
                   "Pterygophora californica") ~ scientific_name)) %>% 
  mutate(spp_interest = case_when(
    is.na(spp_interest) ~ "other species",
    TRUE ~ as.character(spp_interest)
  )) %>% 
  # filter out MAPY
  filter(scientific_name != "Macrocystis pyrifera") %>% 
  # create a new column where total biomass for the whole survey is calculated
  group_by(site) %>% 
  mutate(total_dry = sum(dry_gm2),
         total_wet = sum(wm_gm2)) %>% 
  ungroup() %>% 
  # create a new column where total biomass for the species for the whole survey is calculated
  group_by(site, spp_interest) %>% 
  mutate(total_sp_dry = sum(dry_gm2),
         total_sp_wet = sum(wm_gm2)) %>% 
  ungroup() %>% 
  # create a new column where percent of total biomass per survey for each species is calculated
  mutate(percent_sp_dry = total_sp_dry/total_dry,
         percent_sp_wet = total_sp_wet/total_wet) %>% 
  # double check to make sure the percents worked out...
  select(site, spp_interest, 
         # values that need to be unique: 
         total_sp_dry, total_sp_wet, percent_sp_dry, percent_sp_wet) %>% 
  unique() %>% 
  group_by(site) %>% 
  mutate(total_percent = sum(percent_sp_dry),
         total_percent_wet = sum(percent_sp_wet)) %>% 
  select(spp_interest, site, percent_sp_dry) %>% 
  mutate(percent_sp_dry = round(percent_sp_dry, 4)) %>% 
  mutate(percent_whole_sp_dry = percent_sp_dry*100) %>% 
  select(-percent_sp_dry) %>% 
  mutate(site = str_to_upper(site)) %>% 
  pivot_wider(names_from = "site", values_from = "percent_whole_sp_dry") %>% 
  left_join(., coarse_traits, by = "scientific_name") %>% 
  slice(-2) %>% 
  select(scientific_name, sp_code, BULL, AQUE, AHND, NAPL, IVEE, GOLB, ABUR,
                            MOHK, CARP, SCDI, SCTW)
```

# 3. Urchin density at each site

Plots using `stat_summary` of mean urchin density at each site and a timeseries:

```{r}
urchin_plot <- ggplot(urchins %>% filter(site %in% c("bull", "aque", "napl", "ivee", "mohk", "carp")), aes(x = reorder(site, density), y = density)) +
  stat_summary(aes(fill = site), 
               fun = "mean", geom = "col") +
  stat_summary(aes(group = site),
               fun.data = mean_se, geom = "errorbar", width = 0.5) +
  labs(x = "Site", y = "Mean urchin density (number/m2)",
       title = "Purple urchin density") +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, 35)) +
  theme_bw() +
  theme(legend.position = "none")

urchin_plot

urchins_time_plot <- ggplot(urchins, aes(x = year, y = density)) +
  stat_summary(aes(col = site), 
               fun = "sum", geom = "point", size = 3) +
  stat_summary(aes(col = site),
               fun = "sum", geom = "line") +
  # stat_summary(aes(col = site), 
  #              fun.data = mean_se, geom = "errorbar", width = 0.5) +
  labs(x = "Year", y = "Sum urchin density (number/m2)",
       title = "Purple urchin density through time") +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(-1, 100)) +
  theme_bw()

urchins_time_plot
```

Plot with pairwise comparisons of means:

```{r}
# analysis of variance
urchin_aov <- aov(na.omit(urchins$density) ~ urchins$site)

# Tukey HSD
urchin_tukey <- TukeyHSD(urchin_aov)

# compact letter display
urchin_cld <- multcompLetters4(urchin_aov, urchin_tukey)

# output of compact letter display in data frame
urchin_cld_df <- enframe(urchin_cld[[1]]$Letters)

# summary data frame for plotting
urchin_plot_summary <- urchins %>% 
  # filter(site %in% c("bull", "aque", "ivee", "mohk", "napl", "carp")) %>% 
  group_by(site) %>% 
  summarize(mean = mean(density, na.rm = TRUE),
            se = se(density)) %>% 
  left_join(., urchin_cld_df, by = c("site" = "name")) %>% 
  # mutate(value = fct_relevel(value, "a", "b", "bc", "c")) %>% 
  arrange(value)
  # mutate(site = fct_relevel(site, "napl", "bull", "carp", "aque", "ivee", "mohk"))

# new plot
urchin_plot_new <- ggplot(urchin_plot_summary, aes(x = site, y = mean)) +
  geom_col() +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.5) +
  geom_text(aes(label = value, y = mean + se + 1)) + 
  labs(x = "Site", y = "Mean urchin density (number/m2)",
       title = "Purple urchin density") +
  scale_y_continuous(expand = c(0, 0)) +
  # coord_cartesian(ylim = c(0, 28)) +
  theme_bw()

urchin_plot_new
```


# 4. Sand cover

Plot using `stat_summary`:

```{r}
sand_plot <- ggplot(substrate %>% filter(common_name == "Sand"), aes(x = site, y = percent_cover)) +
  stat_summary(aes(fill = site), 
               fun = "mean", geom = "col") +
  stat_summary(aes(group = site), 
               fun.data = mean_se, geom = "errorbar", width = 0.5) +
  # facet_wrap(~site) +
  labs(x = "Site", y = "Mean sand percent cover (%)",
       title = "Sand cover across sites") +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, 90)) +
  theme_bw() +
  theme(legend.position = "none")

sand_plot

sand_time_plot <- ggplot(substrate %>% filter(common_name == "Sand"), aes(x = year, y = percent_cover)) +
  stat_summary(aes(col = site), 
               fun = "mean", geom = "point", size = 3) +
  stat_summary(aes(col = site),
               fun = "mean", geom = "line") +
  stat_summary(aes(col = site), 
               fun.data = mean_se, geom = "errorbar", width = 0.5) +
  # facet_wrap(~site) +
  labs(x = "Site", y = "Mean sand percent cover (%)",
       title = "Sand cover across sites through time") +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(-1, 105)) +
  theme_bw() 

sand_time_plot
```

In case saving plots is of interest: 

```{r}
plots_together <- urchin_plot / sand_plot

# ggsave(here::here("figures/habitat-characteristics", 
#                   paste("habitat-characteristics-", todays_date, ".jpg", sep = "")), 
#        plots_together)
# 
# plots_together <- urchins_time_plot / sand_time_plot
# 
# ggsave(here::here("figures/habitat-characteristics", 
#                   paste("habitat-characteristics-timeseries", todays_date, ".jpg", sep = "")), 
#        plots_together)
```

Plot with pairwise comparisons of means:

```{r}
# filter out sand
sand <- substrate %>% 
  filter(common_name == "Sand")
  # filter(site %in% c("bull", "aque", "ivee", "mohk", "napl", "carp"))

# analysis of variance
sand_aov <- aov(na.omit(sand$percent_cover) ~ sand$site)

# Tukey HSD
sand_tukey <- TukeyHSD(sand_aov)

# compact letter display
sand_cld <- multcompLetters4(sand_aov, sand_tukey)

# results into data frame
sand_cld_df <- enframe(sand_cld[[1]]$Letters)

# summary data frame
sand_plot_summary <- sand %>% 
  group_by(site) %>% 
  summarize(mean = mean(percent_cover, na.rm = TRUE),
            se = se(percent_cover)) %>% 
  left_join(., sand_cld_df, by = c("site" = "name"))
  # mutate(site = fct_relevel(site, "napl", "bull", "carp", "aque", "ivee", "mohk"))

# new plot
sand_plot_new <- ggplot(sand_plot_summary, aes(x = site, y = mean)) +
  geom_col() +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.5) +
  geom_text(aes(label = value, y = mean + se + 1)) + 
  labs(x = "Site", y = "Mean sand percent cover (%)",
       title = "Sand cover across sites") +
  scale_y_continuous(expand = c(0, 0)) +
  # coord_cartesian(ylim = c(0, 24)) +
  theme_bw()

sand_plot_new
```

# 5. Kelp cover

Plot using `stat_summary`:

```{r}
biomass %>% 
  filter(sp_code == "MAPY") %>% 
  filter(!(site %in% c("scdi", "sctw"))) %>% 
  # replace all -99999 values with 0
  mutate(wm_gm2 = replace(wm_gm2, wm_gm2 < 0, 0)) %>% 
  ggplot(aes(x = reorder(site, wm_gm2), y = wm_gm2)) +
  stat_summary(aes(fill = site), 
               fun = "mean", geom = "col") +
  stat_summary(aes(group = site),
               fun.data = mean_se, geom = "errorbar", width = 0.5) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, 7000)) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(x = "Site", y = "Kelp biomass (wet g/m2)",
       title = "Kelp biomass across sites")

biomass %>% 
  filter(sp_code == "MAPY") %>% 
  filter(site == "carp") %>% 
  # filter(!(site %in% c("scdi", "sctw"))) %>% 
  # replace all -99999 values with 0
  mutate(wm_gm2 = replace(wm_gm2, wm_gm2 < 0, 0)) %>% 
  ggplot(aes(x = year, y = wm_gm2)) +
  stat_summary(aes(col = site), 
               fun = "mean", geom = "point", size = 3) +
  stat_summary(aes(col = site),
               fun = "mean", geom = "line") +
  stat_summary(aes(col = site), 
               fun.data = mean_se, geom = "errorbar", width = 0.5) +
  theme_bw() 
```

Plot with pairwise comparisons:  

**Note:** does it make sense that mean kelp biomass is the same across all sites?

```{r}
mapy <- biomass %>% 
  # filter(!(site %in% c("sctw", "scdi"))) %>% 
  filter(site %in% c("aque", "ivee", "mohk", "napl", "carp")) %>% 
  filter(sp_code == "MAPY") %>% 
  group_by(site, year) %>% 
  summarize(total = sum(dry_gm2)) 

mapy_aov <- aov(na.omit(mapy$total) ~ mapy$site)

mapy_tukey <- TukeyHSD(mapy_aov)

mapy_cld <- multcompLetters4(mapy_aov, mapy_tukey)

mapy_cld_df <- enframe(mapy_cld[[1]]$Letters)

mapy_plot_summary <- mapy %>% 
  group_by(site) %>% 
  summarize(mean = mean(total, na.rm = TRUE),
            se = se(total)) %>% 
  left_join(., mapy_cld_df, by = c("site" = "name")) 

mapy_plot_new <- ggplot(mapy_plot_summary, aes(x = site, y = mean)) +
  geom_col() +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.5) +
  geom_text(aes(label = value, y = mean + se + 70)) + 
  labs(x = "Site", y = "Mean biomass (g dry/m2)",
       title = "MAPY biomass across sites") +
  # scale_y_continuous(expand = c(0, 0)) +
  # coord_cartesian(ylim = c(0, 1600)) +
  theme_bw()

mapy_plot_new
```

# 6. Understory biomass

```{r}
macrobio <- biomass %>% 
  filter(site %in% c("aque", "ivee", "mohk", "napl", "carp")) %>% 
  filter(sp_code != "MAPY") %>% 
  group_by(site, year) %>% 
  summarize(total = sum(dry_gm2))

macrobio_aov <- aov(macrobio$total ~ macrobio$site)

macrobio_tukey <- TukeyHSD(macrobio_aov)

macrobio_cld <- multcompLetters4(macrobio_aov, macrobio_tukey)

macrobio_cld_df <- enframe(macrobio_cld[[1]]$Letters)

macrobio_plot_summary <- macrobio %>% 
  group_by(site) %>% 
  summarize(mean = mean(total, na.rm = TRUE),
            se = sd(total)/sqrt(length(total))) %>% 
  left_join(., macrobio_cld_df, by = c("site" = "name")) 
  # mutate(site = fct_relevel(site, "napl", "carp", "aque", "ivee", "mohk"))

macrobio_plot_new <- ggplot(macrobio_plot_summary, aes(x = site, y = mean)) +
  geom_col() +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.5) +
  geom_text(aes(label = value, y = mean + se + 50)) + 
  labs(x = "Site", y = "Mean biomass (g dry/m2)",
       title = "Understory macroalgal biomass across sites") +
  scale_y_continuous(expand = c(0, 0)) +
  # coord_cartesian(ylim = c(0, 1600)) +
  theme_bw()

macrobio_plot_new

# all algae

all_macro <- biomass %>% 
  filter(site %in% c("aque", "ivee", "mohk", "napl", "carp")) %>% 
  group_by(site, year) %>% 
  summarize(total = sum(dry_gm2))

all_macro_aov <- aov(all_macro$total ~ all_macro$site)

all_macro_tukey <- TukeyHSD(all_macro_aov)

all_macro_cld <- multcompLetters4(all_macro_aov, all_macro_tukey)

all_macro_cld_df <- enframe(all_macro_cld[[1]]$Letters)

all_macro_plot_summary <- all_macro %>% 
  group_by(site) %>% 
  summarize(mean = mean(total, na.rm = TRUE),
            se = sd(total)/sqrt(length(total))) %>% 
  left_join(., all_macro_cld_df, by = c("site" = "name")) 
  # mutate(site = fct_relevel(site, "napl", "carp", "aque", "ivee", "mohk"))

all_macro_plot_new <- ggplot(all_macro_plot_summary, aes(x = site, y = mean)) +
  geom_col() +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.5) +
  geom_text(aes(label = value, y = mean + se + 80)) + 
  labs(x = "Site", y = "Mean biomass (g dry/m2)",
       title = "Whole community biomass across sites") +
  scale_y_continuous(expand = c(0, 0)) +
  # coord_cartesian(ylim = c(0, 4500)) +
  theme_bw()

all_macro_plot_new
```

# 7. Community metrics

## a. wrangling

Going to make a site by species matrix for each sampling date. There are a couple of anomalies: CARP_2000-09-08 and NAPL_2000-09-18 don't have any observations, and AQUE_2019-07-23 is missing the sand cover data because sand cover was taken on 2019-07-24. Going to fix that later.   

**Note:** as of 2022-03-04, the code to create `community_matrix` and `community_metadata` are in the `00_set-up.R` script.

## b. Species richness/diversity

### i. Richness

```{r}
sp_rich <- specnumber(community_matrix) %>% 
  enframe() %>% 
  left_join(., community_metadata, by = c("name" = "sample_ID")) %>% 
  rename(richness = value)

sp_rich_aov <- aov(sp_rich$richness ~ sp_rich$site)

sp_rich_tukey <- TukeyHSD(sp_rich_aov)

sp_rich_cld <- multcompLetters4(sp_rich_aov, sp_rich_tukey)

sp_rich_cld_df <- enframe(sp_rich_cld[[1]]$Letters)

sp_rich_plot_summary <- sp_rich %>% 
  group_by(site) %>% 
  summarize(mean = mean(richness, na.rm = TRUE),
            se = sd(richness)/sqrt(length(richness))) %>% 
  left_join(., sp_rich_cld_df, by = c("site" = "name")) %>% 
  mutate(site = fct_relevel(site, "napl", "carp", "aque", "ivee", "mohk"))

sp_rich_plot_new <- ggplot(sp_rich_plot_summary, aes(x = site, y = mean)) +
  geom_col() +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.5) +
  geom_text(aes(label = value, y = mean + se + 1)) + 
  labs(x = "Site", y = "Mean species richness",
       title = "Understory macroalgal species richness across sites") +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, 25)) +
  theme_bw()

sp_rich_plot_new
```

### ii. Diversity
```{r}
sp_div <- diversity(community_matrix, index = "shannon") %>% 
  enframe() %>% 
  left_join(., community_metadata, by = c("name" = "sample_ID")) %>% 
  rename(shandiv = value)

sp_div_aov <- aov(sp_div$shandiv ~ sp_div$site)

sp_div_tukey <- TukeyHSD(sp_div_aov)

sp_div_cld <- multcompLetters4(sp_div_aov, sp_div_tukey)

sp_div_cld_df <- enframe(sp_div_cld[[1]]$Letters)

sp_div_plot_summary <- sp_div %>% 
  group_by(site) %>% 
  summarize(mean = mean(shandiv, na.rm = TRUE),
            se = sd(shandiv)/sqrt(length(shandiv))) %>% 
  left_join(., sp_div_cld_df, by = c("site" = "name")) %>% 
  mutate(site = fct_relevel(site, "napl", "carp", "aque", "ivee", "mohk"))

sp_div_plot_new <- ggplot(sp_div_plot_summary, aes(x = site, y = mean)) +
  geom_col() +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.5) +
  geom_text(aes(label = value, y = mean + se + 0.2)) + 
  labs(x = "Site", y = "Mean Shannon diversity",
       title = "Understory macroalgal Shannon diversity across sites") +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, 2.75)) +
  theme_bw()

sp_div_plot_new

sp_div %>% 
  ggplot(aes(x = site, y = shandiv)) +
  geom_violin() +
  geom_point() +
  geom_text(data = sp_div_cld_df, aes(x = name, y = 3, label = value))
```


## c. SIMPER analysis

Wanted to know which species significantly contributed to differences in community composition.

```{r}
sim <- with(community_metadata, simper(community_matrix, site)) 
head(summary(sim))
```

## d. NMDS 

### i. Jaccard (presence/absence)

```{r}
community_mds_jacc <- metaMDS(community_presabs, distance = "jaccard")

community_dist_jacc <- vegdist(community_presabs, method = "jaccard")

stressplot(community_mds_jacc)

community_mds_sites_jacc <- scores(community_mds_jacc, display = "sites") %>% 
  as.data.frame() %>% 
  rownames_to_column("sample_ID") %>% 
  left_join(., community_metadata, by = "sample_ID") 

community_mds_spp_jacc <- scores(community_mds_jacc, display = "species") %>% 
  as.data.frame() %>% 
  rownames_to_column("sp_code") %>% 
  left_join(., coarse_traits, by = "sp_code") %>% 
  filter(sp_code %in% algae_proposal)

perm <- adonis(community_dist_jacc ~ site, data = community_metadata)

perm

pairwise.results <- pairwise.adonis(community_presabs, factors = community_metadata$site, sim.method = "jaccard")

disper <- betadisper(community_dist_jacc, community_metadata$site)

anova(disper)

TukeyHSD(disper)

fit_jacc <- community_metadata %>% 
  select(sample_ID, site, urchin_density, mean_sand) %>% 
  mutate(across(.cols = 3:4, ~replace(., is.na(.), 0))) %>% 
  envfit(community_mds_jacc, ., perm = 999)

fit_vect_jacc <- scores(fit, display = "vectors") %>% 
  as.data.frame()

nmds_plot_jacc <- ggplot() +
  # x and y axes
  geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +
  # points for sites
  geom_point(data = community_mds_sites_jacc, aes(x = NMDS1, y = NMDS2, color = site)) +
  stat_ellipse(data = community_mds_sites_jacc, aes(x = NMDS1, y = NMDS2, color = site), level = 0.95, size = 1) +
  # # arrows for species
  # geom_segment(data = community_mds_spp_jacc, aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), arrow = arrow(length = unit(0.2, "cm")), color = "blue") +
  # geom_text(data = community_mds_spp_jacc, aes(x = NMDS1, y = NMDS2, label = scientific_name), color = "blue") +
  # # arrows for environmental characteristics (vector fitting)
  # geom_segment(data = fit_vect_jacc, aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), arrow = arrow(length = unit(0.2, "cm"))) +
  # geom_text(data = fit_vect_jacc, aes(x = NMDS1, y = NMDS2, label = rownames(fit.vect))) +
  # theme
  theme_bw() +
  guides(colour = guide_legend(nrow = 5)) +
  theme(legend.position = "bottom", legend.box = "vertical", legend.margin = margin()) 

nmds_plot_jacc
```


### ii. Bray-Curtis

```{r}
community_mds_bray <- metaMDS(community_matrix, distance = "bray")

community_dist_bray <- vegdist(community_matrix, method = "bray")

stressplot(community_mds_bray)

community_mds_sites_bray <- scores(community_mds_bray, display = "sites") %>% 
  as.data.frame() %>% 
  rownames_to_column("sample_ID") %>% 
  left_join(., community_metadata, by = "sample_ID") 

community_mds_spp_bray <- scores(community_mds_bray, display = "species") %>% 
  as.data.frame() %>% 
  rownames_to_column("sp_code") %>% 
  left_join(., coarse_traits, by = "sp_code") %>% 
  filter(sp_code %in% algae_proposal)

perm <- adonis(community_dist_bray ~ site, data = community_metadata)

perm

pairwise.results <- pairwise.adonis(community_matrix, factors = community_metadata$site, sim.method = "bray")

disper <- betadisper(community_dist_bray, community_metadata$site)

anova(disper)

TukeyHSD(disper)

fit <- community_metadata %>% 
  select(sample_ID, site, urchin_density, mean_sand) %>% 
  mutate(across(.cols = 3:4, ~replace(., is.na(.), 0))) %>% 
  envfit(community_mds_bray, ., perm = 999)

fit_vect_bray <- scores(fit, display = "vectors") %>% 
  as.data.frame()

nmds_plot_bray <- ggplot() +
  geom_point(data = community_mds_sites_bray, aes(x = NMDS1, y = NMDS2, color = site)) +
  # scale_color_manual(values = pal) +
  geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +
  stat_ellipse(data = community_mds_sites_bray, aes(x = NMDS1, y = NMDS2, color = site), level = 0.95, size = 1) +
  # geom_segment(data = community_mds_scores_spp, aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), arrow = arrow(length = unit(0.2, "cm"))) +
  # geom_text(data = community_mds_scores_spp, aes(x = NMDS1, y = NMDS2, label = scientific_name)) +
  geom_segment(data = fit_vect_bray, aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), arrow = arrow(length = unit(0.2, "cm"))) +
  geom_text(data = fit_vect_bray, aes(x = NMDS1, y = NMDS2, label = rownames(fit.vect))) +
  theme_bw() +
  guides(colour = guide_legend(nrow = 5)) +
  theme(legend.position = "bottom", legend.box = "vertical", legend.margin = margin()) 

nmds_plot_bray
```

## e. RDA (in proposal)

```{r}
community_rda <- rda(community_matrix ~ site + urchin_sqrt + sand_sqrt + kelp_sqrt, data = community_metadata)
summary(community_rda)

varpart(community_matrix, ~ site + urchin_sqrt + sand_sqrt + kelp_sqrt, data = community_metadata)

# RDA1
3.158e-01 
# RDA2:
2.277e-02
# RDA 3: 
1.563e-02 
```

# 8. Trait calculations

## a. proportion of traits

"Mean proportion of biomass" for each species per trait across sampling dates: matrix where row is sampling date, and columns are traits spread across, and cells are total biomass that can be "assigned" to that trait. The `prop_biomass` data frame is already set up.

### i. growth form

Going to test this out with a couple of sites just to make sure the math is right.

```{r}
prop_traits_test <- prop_biomass %>% 
  filter(sample_ID %in% c("CARP_2000-09-08", "BULL_2016-07-25")) %>% 
  select(sample_ID, site, year, sp_code, growth_form,
         total_sp_dry, total_sp_wet, percent_sp_dry, percent_sp_wet) %>% 
  # create a new column where total biomass for the whole survey is calculated
  group_by(site, year) %>% 
  mutate(total_dry = sum(total_sp_dry),
         total_wet = sum(total_sp_wet)) %>% 
  ungroup() %>% 
  # calculate biomass within each growth form
  group_by(site, year, growth_form) %>% 
  mutate(total_gf_dry = sum(total_sp_dry),
         total_gf_wet = sum(total_sp_wet),
         percent_gf_dry = total_gf_dry/total_dry,
         percent_gf_wet = total_gf_wet/total_wet) %>% 
  ungroup() %>% 
  # check to make sure that if you were to sum all the percents for growth forms for each sampling ID, it would add to 1
  group_by(site, year) %>% 
  mutate(sum_percent_dry = sum(unique(percent_gf_dry))) %>% 
  ungroup() %>% 
  # make the data frame wider
  select(sample_ID, growth_form, percent_gf_dry) %>% 
  unique() %>% 
  pivot_wider(names_from = "growth_form", values_from = "percent_gf_dry")
```

Now that it seems ok, going to do the rest of the data set.

```{r}
prop_traits <- prop_biomass %>% 
  # filter(site %in% c("bull", "aque", "ahnd", "napl", "ivee", "golb", 
  #          "abur", "mohk", "carp")) %>% 
  select(sample_ID, site, year, sp_code, growth_form,
         total_sp_dry, total_sp_wet, percent_sp_dry, percent_sp_wet) %>% 
  # create a new column where total biomass for the whole survey is calculated
  group_by(site, year) %>% 
  mutate(total_dry = sum(total_sp_dry),
         total_wet = sum(total_sp_wet)) %>% 
  ungroup() %>% 
  # calculate biomass within each growth form
  group_by(site, year, growth_form) %>% 
  mutate(total_gf_dry = sum(total_sp_dry),
         total_gf_wet = sum(total_sp_wet),
         percent_gf_dry = total_gf_dry/total_dry,
         percent_gf_wet = total_gf_wet/total_wet) %>% 
  ungroup() %>% 
  # make the data frame wider
  select(sample_ID, growth_form, percent_gf_dry) %>% 
  unique() %>% 
  pivot_wider(names_from = "growth_form", values_from = "percent_gf_dry")
```

GLM???

```{r}
df <- biomass %>% 
  # filter sites of interest
  filter(site %in% c("bull", "aque", "ahnd", "napl", "ivee", "golb", 
           "abur", "mohk", "carp")) %>% 
  # only include understory species
  filter(sp_code != "MAPY") %>% 
  # join with trait data
  left_join(., coarse_traits, by = "sp_code") %>% 
  group_by(sample_ID, site, year, date, growth_form) %>% 
  summarize(mean_gf_dry = mean(dry_gm2, na.omit = TRUE),
         mean_gf_wet = mean(wm_gm2, na.omit = TRUE),
         total_gf_dry = sum(dry_gm2),
         total_gf_wet = sum(wm_gm2)) %>% 
  ungroup() %>% 
  # create a new column where total biomass for the whole survey is calculated
  group_by(site, year) %>% 
  mutate(total_dry = sum(total_gf_dry),
         total_wet = sum(total_gf_wet)) %>% 
  ungroup() %>% 
  # create new column where percent growth form biomass is calculated for each survey
  mutate(percent_gf_dry = total_gf_dry/total_dry,
         percent_gf_wet = total_gf_wet/total_wet) %>% 
  # join with metadata data frame
  left_join(., community_metadata, by = "sample_ID") %>% 
  filter(growth_form == "leathery_macrophyte") %>% 
  # filter(total_gf_dry > 0) %>% 
  drop_na()

ggplot(df, aes(x = urchin_density, y = mean_gf_dry)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) 

ggplot(df, aes(x = mean_sand, y = mean_gf_dry)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) 

ggplot(df %>% filter(mean_gf_dry > 0), aes(x = mean_gf_dry)) +
  geom_histogram()

# GLM with tweedie distribution for leathery macrophytes
model1 <- glmmTMB(mean_gf_dry ~ urchin_density + mean_sand + urchin_density*mean_sand + (1|site.x) + (1|year.x), 
              data = df, family = tweedie(link = "log"))

summary(model1)

simul.resid <- simulateResiduals(model1)

plot(simul.resid)


testZeroInflation(model1)
# no zero inflation - which makes sense, because it's a tweedie distribution
testDispersion(simul.resid)
# no over dispersion

# GLM with gamma distribution and a log link function, but without the interaction between urchins and sand
model2 <- glmmTMB(total_gf_dry ~ urchin_density + mean_sand + (1|site.x) + (1|year.x), 
              data = df, family = Gamma(link = "log"))

summary(model2)

AIC(model1, model2)

predict.glm(model1, type = "response")

# GLM with beta
model1.prop <- glmmTMB(percent_gf_dry ~ urchin_density + mean_sand + mean_kelp + (1|site.x) + (1|year.x),
                       data = df, family = beta_family(link = "logit"))

summary(model1.prop)

emmeans(model1, "urchin_density")
```

# 9. Community-weighted means

These are calculated using the `functcomp` function in the `FD` package. The function requires a matrix with functional traits and a matrix of species abundances (or presence-absence). In the trait matrix, rows are species and columns are traits. In the species matrix, rows are samples and columns are species.  

The first step is to get the `coarse_traits` data frame into a functional form (lol).

```{r}
trait_matrix <- coarse_traits %>% 
  select(sp_code, growth_morph, growth_form, pigment_type, 
         height_class, longevity, posture, branching_yn) %>% 
  filter(sp_code != "MAPY") %>% 
  column_to_rownames("sp_code")

# making sure the rownames in the trait matrix == columns in the community matrix
rownames(trait_matrix) == colnames(community_matrix)
```

Then the community matrix has to be a matrix, not a data frame (why?).
```{r}
comm_matrix <- as.matrix(community_matrix)
```


Then, you're ready to calculate community-weighted means.

```{r}
CWM_cat <- functcomp(trait_matrix, comm_matrix, 
                  CWM.type = c("all")) %>% 
  rownames_to_column("sample_ID") %>% 
  left_join(., community_metadata, by = "sample_ID") %>% 
  # to get the glmm to work!!!!!!
  mutate(across(.cols = everything(), .fns = ~ replace(., . == 1, 0.999)))
```

For categorical traits, the output is the exact same as the `prop_traits` data frame up top - each cell is the proportion of the community biomass that is attributed to a specific category (for example, within `growth_morph`, at CARP in 2000, there is 92% `solitary` and 8 % `aggregated`).

Now to plot?

```{r}
lm_plot <- ggplot(CWM_cat, 
                  aes(x = urchin_density, y = growth_form_leathery_macrophyte)) +
  # scale_x_continuous(expand = c(0, 0)) +
  # scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(-0.001, 1)) +
  geom_point(aes(col = site)) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Urchin density", y = "Proportion biomass (dry g/m2)",
       title = "Proportion leathery macrophytes")

lm_plot

model_lm <- glmmTMB(growth_form_leathery_macrophyte ~ urchin_density + (1|site) + (1|year), 
              data = CWM_cat, ziformula = ~1, family = beta_family(link = "logit"))

model_lm

summary(model_lm)

model_lm_resid <- simulateResiduals(model_lm)

plot(model_lm_resid)

temp <- ggpredict(model_lm, terms = "urchin_density")

# predict(model_lm)

plot(temp, add.data = TRUE)
```


```{r}
ggplot(CWM_cat, aes(x = urchin_density, y = mean_sand)) +
  geom_point() 

cor.test(CWM_cat$urchin_density, CWM_cat$mean_sand)
```




```{r include = FALSE}
# Use `rmarkdown::render(here::here("code", "04_benthic-data.Rmd"))` in the console to knit the document.
```

