---
title: "Proposal-Box2"
author: "Bart DiFiore"
date: "4/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In the first box, I demonstrate how variation in a trait can cause the same two species to interact differently in different contexts. 

Here, I would like to demonstrate that an alternative way for the same species to interact differently is based on historical context.

I would like to build a toy discrete time model with three interacting species. Species 1 competes very strongly with species 2, such that species 1 and 3 coexist. I will test how historic variability in Species 1 can lead to different interaction strengths between the three species compared to simulations with low historic variability.

Species 1 represents a competitive dominant. It is fast growing and competitive but highly suseptible to disturbance. 

Species 2 is a competitive inferior. It is slower growing than 1, not very suseptible to disturbance, and negatively affected by species 1. 

Species 3 is a slow growing species that can coexist with species 1, but is outcompeted by species 2.

$$
\frac{dN_i}{dt} = r_iN_i(\frac{k_i - N_i - \alpha_{ij}N_j}{k_i}) 
$$

All species grow to their carrying capacity in the absence of competitors.

```{r}
# Set some parameter values

# library(tidyverse)
k0 <- 1000
u0 <- 100
i0 <- 100
r.k <- 0.3
r.u <- 0.2
r.i <- 0.1
k.k <- 10000
k.u <- 5000
k.i <- 3000
alpha.ku <- 0.1
alpha.ki <- 0.01
alpha.uk <- 0.5
alpha.ui <- 0.1
alpha.ik <- 0.01
alpha.iu <- 0.1
recruit <- 500



tset <- seq(from = 0, to = 2000, length.out = 200000)
k.simu <- NaN*tset; k.simu[1] <- k0
u.simu <- NaN*tset; u.simu[1] <- u0 
i.simu <- NaN*tset; i.simu[1] <- i0 
dist <- rbinom(n = length(tset), size = 1, prob = 0.00001)

for(i in 2:length(tset)){
  dt <- tset[i]-tset[i-1]

    if (dist[i] > 0){
      K <- k.simu[i-1]*0.1
      U <- u.simu[i-1]*0.9
      I <- i.simu[i-1]*0.9
  } else{
    
    K <- k.simu[i-1]
    U <- u.simu[i-1]
    I <- i.simu[i-1]
  }
    
    dK <- (r.k*K*((k.k - K - alpha.ku*U - alpha.ki*I)/k.k) + recruit) * dt
    dU <- (r.u*U*((k.u - U - alpha.uk*K - alpha.ui*I)/k.u) + recruit) * dt
    dI <- (r.i*I*((k.i - I - alpha.ik*K - alpha.iu*U)/k.i) + recruit) * dt
    
    k.simu[i] <- K + dK
    u.simu[i] <- U + dU
    i.simu[i] <- I + dI
}

df.high <- data.frame(K = k.simu, U = u.simu, I = i.simu, t = tset) %>% pivot_longer(cols = c(K,U,I))
df.low <- data.frame(K = k.simu, U = u.simu, I = i.simu, t = tset) %>% pivot_longer(cols = c(K,U,I))

p2 <- ggplot(df.low, aes(x= t, y = value))+
  geom_line(aes(color = name))+
  scale_y_continuous(limits = c(0, 15000)) +
  theme_classic()

p2

```

```{r, include = T}

# Set some parameter values
parameters <- list(
k0 = 1000,
u0 = 100,
i0 = 100,
r.k = 0.3,
r.u = 0.2,
r.i = 0.1,
k.k = 10000,
k.u = 5000,
k.i = 3000,
alpha.ku = 0.1,
alpha.ki = 0.01,
alpha.uk = 0.5,
alpha.ui = 0.1,
alpha.ik = 0.01,
alpha.iu = 0.1,
recruit = 0.001
)

project <- function(params, end.t = 2000, prob = 0.0001){

tset <- seq(from = 0, to = end.t, by = 0.01)
k.simu <- NaN*tset; k.simu[1] <- params$k0
u.simu <- NaN*tset; u.simu[1] <- params$u0 
i.simu <- NaN*tset; i.simu[1] <- params$i0 
dist <- rbinom(n = length(end.t), size = 1, prob = prob)


for(i in 2:length(tset)){
  dt <- tset[i]-tset[i-1]

    if (dist[i] > 0){
      K <- k.simu[i-1]*0.1
      U <- u.simu[i-1]*0.9
      I <- i.simu[i-1]*0.9
  } else{
    
    K <- k.simu[i-1]
    U <- u.simu[i-1]
    I <- i.simu[i-1]
  }
    
    dK <- with(params, (r.k*K*((k.k - K - alpha.ku*U - alpha.ki*I)/k.k) + recruit)) * dt
    dU <- with(params, (r.u*U*((k.u - U - alpha.uk*K - alpha.ui*I)/k.u) + recruit)) * dt
    dI <- with(params, (r.i*I*((k.i - I - alpha.ik*K - alpha.iu*U)/k.i) + recruit)) * dt
    
    k.simu[i] <- K + dK
    u.simu[i] <- U + dU
    i.simu[i] <- I + dI
}
 data.frame(K = k.simu, U = u.simu, I = i.simu, t = tset) %>% pivot_longer(cols = c(K,U,I))
}

df.low = project(params = parameters, end.t = 2000, prob = 0.0001)
df.high = project(params = parameters, end.t = 2000, prob = 0.001)

p1 <- ggplot(df.low, aes(x= t, y = value))+
  geom_line(aes(color = name))+
  theme_classic()

p2 <- ggplot(df.high, aes(x= t, y = value))+
  geom_line(aes(color = name))+
  theme_classic()

cowplot::plot_grid(p1, p2)

```



































